Biofouling

### 2023-07-11 smelt_2019_allseasons/Biofouling.Rmd

---
title: "Biofouling"
author: "Nicole Kwan"
date: "2023-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

---
title: "Biofouling"
author: "Nicole Kwan"
date: "2023-06-09"
output: html_document
editor_options: 
  chunk_output_type: console
---

# load packages
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(ggpubr)
```

# bring in & prep biofouling data
```{r}
biofoul <- read_csv("smelt_2019_allseasons/data_raw/Invert.data.csv")
unique(biofoul$Taxa)

#match ID level to diet data
biofoul$Taxa2 <- biofoul$Taxa
biofoul[biofoul$Taxa %in% c("Corphiidae", "Gammaridae"),]$Taxa2 <- "Amphipod"

#sum across location
biofoul2<- biofoul %>% 
  group_by(Taxa2, Location) %>% 
  summarize("Count.sum"= sum(Count))

#rename columns
biofoul3 <- biofoul2 %>% rename(biofoul = "Count.sum")
```

# bring in & prep diet data
```{r}
diet <- read_csv("smelt_2019_allseasons/data_clean/Diet.proportions.csv")
unique(diet$Taxa)

#match up taxa names to work with biofoul data
diet$Taxa2 <- diet$Taxa
diet[diet$Taxa %in% c("Cyclopoid", "Calanoid"),]$Taxa2 <- "Copepod"
diet[diet$Taxa %in% c("Ostracod", "Cladocera"),]$Taxa2 <- "Other"

#sum across location
diet2<- diet %>% 
  group_by(Taxa2, Location) %>% 
  summarize("Count.sum"= sum(Count))

unique(diet2$Location)

#filter to just fall at core sites
diet3 <- diet2 %>% filter(Location == "RV.Fall" | Location == "SM.Fall" | Location == "YB.Fall")

#remove empty
diet4<-diet3[diet3$Taxa2 != "empty",]

#change Location to drop season
diet4$Location[diet4$Location=="RV.Fall"] <- "RV"
diet4$Location[diet4$Location =="YB.Fall"] <- "YB"
diet4$Location[diet4$Location =="SM.Fall"] <- "SM"

#rename columns
diet4 <- diet4 %>% rename(diet = "Count.sum")
```

# bring in & prep zooplankton data & merge all datasets
```{r}
zoop <- read_csv("smelt_2019_allseasons/data_clean/Zoop.catch.csv")
unique(zoop$Taxa)

#filter zoop dataset to final day of sampling 
zoop2 <- filter(zoop, Date == "2019-11-04")

#remove microzoop
zoop2 <- zoop2[!(zoop2$Taxa=="Microzoo"),]

#match up taxa names to work with biofoul data
zoop2$Taxa2 <- zoop2$Taxa
zoop2[zoop2$Taxa %in% c("Cyclopoid", "Calanoid"),]$Taxa2 <- "Copepod"
zoop2[zoop2$Taxa %in% c("Ostracod", "Cladocera"),]$Taxa2 <- "Other"

#sum across new taxa
zoop3<- zoop2 %>% 
  group_by(Taxa2, Location) %>% 
  summarize("sum.indm3"= sum(ind_m3))

#rename columns
zoop4 <- zoop3 %>% rename(zoop = "sum.indm3")

#combine
all <- merge(x=diet4,y=zoop4,
        by=c("Location","Taxa2"), all=TRUE)

all2 <- merge(x=all,y=biofoul3,
        by=c("Location","Taxa2"), all=TRUE)

#swap NAs with 0
all2[is.na(all2)] <- 0
```

# plot
```{r}
b<-ggplot(data=all2, aes(x=Location, y =biofoul, fill=Taxa2)) +
    geom_bar(position="fill", stat="identity", colour="black") +
  theme_bw() +
  viridis::scale_fill_viridis(discrete = TRUE, option = "plasma") +
  theme(legend.position="right", axis.title.x = element_blank())+
  labs(y = "Biofoul samples")

z<-ggplot(data=all2, aes(x=Location, y =zoop, fill=Taxa2)) +
    geom_bar(position="fill", stat="identity", colour="black") +
  theme_bw() +
  viridis::scale_fill_viridis(discrete = TRUE, option = "plasma") +
  theme(legend.position="right", axis.title.x = element_blank())+
  labs(y = "Zooplankton samples") #final day, no microzoop

d<-ggplot(data=all2, aes(x=Location, y =diet, fill=Taxa2)) +
    geom_bar(position="fill", stat="identity", colour="black") +
  theme_bw() +
  viridis::scale_fill_viridis(discrete = TRUE, option = "plasma") +
  theme(legend.position="right", axis.title.x = element_blank())+
  labs(y = "Stomach contents")

multiplot <- ggarrange(b,z,d, nrow=3, ncol=1)
multiplot
```

# cross-ref what that big chunk of "other" is in the zoop data
```{r}
#sum across new taxa
zoop.only<- zoop2 %>% 
  group_by(Taxa, Location) %>% 
  summarize("sum.indm3"= sum(ind_m3))

ggplot(data=zoop.only, aes(x=Location, y =sum.indm3, fill=Taxa)) +
  geom_bar(position="fill", stat="identity", colour="black") +
  theme_bw() +
  viridis::scale_fill_viridis(discrete = TRUE, option = "plasma") +
  theme(legend.position="right", axis.title.x = element_blank())+
  labs(y = "Zooplankton samples - final day")
```
